---
title: "GDAC Capstone Project: Bellabeat"
author: "Alifia Ganjaraharja"
date: "2025-12-23"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set-directory}
# Create a code chunk and set your working directory
setwd("~/Downloads/CapstoneProject-02")
```

## **INTRODUCTION**

**OVERVIEW OF THE BELLABEAT COMPANY**

Bellabeat is the go-to wellness brand for women, offering an ecosystem
of products and services focused on women's health. Founded by Urša
Sršen and Sando Mur, Bellabeat aims to analyze data that could help
unlock new opportunities and gain valuable marketing strategies.

**STAKEHOLDER**

The stakeholder include:

1\. Urška Sršen, Bellabeat co-founder and Chief Executive Officer

2\. Sando Mur, Mathematician and Bellabeat's Co-founder

3\. Bellabeat's Marketing Analytics Team

**Business Task**

Bellabeat aims to analyze the usage data from one of its products to
gain insights and make high-level recommendations that will inform its
marketing strategy.

**Questions for Analysis**

1.  What are some trends found in smart device usage?
2.  How could these trends affect Bellabeat customers?
3.  How could these trends impact Bellabeat's marketing strategy

## **DATA PREPARATION**

**Data Source**:

The **FitBit Fitness Tracker Data** dataset by **Mobius** under the
license **CCO: Public Domain** was used. This dataset was generated by
respondents to a distributed survey via Amazon Mechanical Turk between
03.12.206 - 05.12.2016. Around 30 eligible FitBit users consented to the
submission of personal tracker data, including minute-level output for
physical activity, heart rate, and sleep monitoring.

**Sorting & Filtering**

My analysis is focused on trends in the usage of the app, which is why
my analysis will be focused on user engagament.

## Load Packages

```{r setup-01, message=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(dplyr)
library(lubridate)
library(tidyr)
```

## Importing the Datasets

```{r load-data}
DailyActivity_01 <- read_csv("archive/mturkfitbit_export_3.12.16-4.11.16/Fitabase Data 3.12.16-4.11.16/dailyActivity_merged.csv")
DailyActivity_02 <- read_csv("archive/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/dailyActivity_merged.csv")
DailySleep <- read_csv("archive/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/sleepDay_merged.csv")
HourlySteps_01 <- read_csv("archive/mturkfitbit_export_3.12.16-4.11.16/Fitabase Data 3.12.16-4.11.16/hourlySteps_merged.csv")
HourlySteps_02 <- read_csv("archive/mturkfitbit_export_4.12.16-5.12.16/Fitabase Data 4.12.16-5.12.16/hourlySteps_merged.csv")
```

## Preview Data

To prove that the file has successfully loaded, here are the first few
rows of one of the dataframe, which is DailyActvity_01.

```{r preview-data}
head(DailyActivity_01)
```

### Check Sample Size

Next, let's take a look at and check the sample size of each dataframe.

```{r check-sample-size}
# Checking distinct values to find out the sample size
n_distinct(DailyActivity_01$Id)
n_distinct(DailyActivity_02$Id)
n_distinct(DailySleep$Id)
n_distinct(HourlySteps_01$Id)
n_distinct(HourlySteps_02$Id)
```

From this information, it is evident that there are between 33 and 35
partcipants in most dataframes; however, the DailySleep dataframe
contains only 24 users.

## **DATA CLEANING**

### Merge Datasets

DailyActivity_01 and DailyActivity_02 are essentially identical, except
for their different time periods. Therefore, I need to merge them into a
single dataframe named 'DailyActivity'. As well as DailyActivity,
HourlySteps_01 & HourlySteps_02 are identical,except for their different
time periods. Therefore, I need to merge them into a single dataframe
too named 'HourlySteps'.

```{r merge-data}
# Merge DailyActivity_01 and DailyActivity_02 together
DailyActivity <- rbind(DailyActivity_01, DailyActivity_02)

# Merge HourlyStep_01 and HourlyStep_02 together
HourlySteps <- rbind(HourlySteps_01, HourlySteps_02)
```

I discovered an issue with the data types of the **'ActivityDate'**
columns across the DailyActivity datraframes, as well as
**'ActivityHour'** columns across the HourlySteps dataframes, and
**'SleepDay'** column across the DailySleep dataframe. Before proceeding
with further analysis, I need to convert them from characters to a date
format.

### Create Clean Data Frames

Before proceeding, I created a duplicate dataframe for every dataframe
named beginning with **Clean_DataframeName** to ensure the original data
remains unchanged.

```{r create-clean-df for DailyActivity}
# Create a new data frame for DailyActivity
Clean_DailyActivity <- DailyActivity

# Verify
head(Clean_DailyActivity)
```

### Format Dates

```{r format-dates for ActivityDate}
library(lubridate)

# Convert ActivityDate to proper date format
Clean_DailyActivity$ActivityDate <- mdy(Clean_DailyActivity$ActivityDate)

# Check the results
head(Clean_DailyActivity$ActivityDate)
class(Clean_DailyActivity$ActivityDate)
```

Now the data type of the *'ActivityDate'* column in the
Clean_DailyActivity frame is changed to Date.

Next, it's time to change the data type of the *'ActivityHour'* column
in the HourlySteps dataframe, but first let's create a new dataframe
named 'Clean_HourlySteps' so that the original dataframe remains the
same.

```{r create-clean-df for HourlySteps }
# Create a new data frame for HourlySteps
Clean_HourlySteps <- HourlySteps

# Verify
head(Clean_HourlySteps)
```

Now, change the format to Date.

```{r format-dates for ActivityHous}
# Convert ActivityHour to proper date format
Clean_HourlySteps <- Clean_HourlySteps %>%
  mutate(
    ActivityHour = mdy_hms(ActivityHour),  # converts "3/12/2016 10:00:00 AM"
    hour = hour(ActivityHour)
  )

# Check the results
head(Clean_HourlySteps$ActivityHour)
class(Clean_HourlySteps$ActivityHour)
```

Now the data type of the *'ActivityHour'* column in the
Clean_HourlySteps data frame is changed to POSIXct.

Last, I need to change the data type of 'SleepDay' column from character
to date, but same like two other dataframe, I need to create a new
dataframe so that the original dataframe remain the same vallues.

```{r create-clean-df for DailySleep}
# Create a new data frame for DailySleep
Clean_DailySleep <- DailySleep

# Verify
head(Clean_DailySleep)
```

Now, change the format to Date.

```{r format-dates for SleepDay}
# Convert SleepDay to proper date format
Clean_DailySleep <- Clean_DailySleep %>%
  mutate(
    SleepDay = mdy_hms(SleepDay),  # converts "3/12/2016 10:00:00 AM"
    hour = hour(SleepDay)
  )

# Check the results
head(Clean_DailySleep$SleepDay)
class(Clean_DailySleep$SleepDay)
```

Now the data type of the *'SleepDay'* column in the Clean_DailySleep
data frame is changed to POSIXct.

After that, I need to find out if there are duplicate users within these
3 dataframes (Clean_DailyActivity, Clean_DailySleep, and
Clean_HourlySteps)

### Check Duplicate

```{r check duplicate on DailyActivity}
#Checking any duplicate based on user + date on DailyActivity dataframe
sum(duplicated(Clean_DailyActivity[, c("Id", "ActivityDate")]))
```

Turned out that 24 duplicate rows are appearing on the DailyActivity /
Clean_DailyActivity dataframe with different values. The next step I
take is to remove the row with the lowest value and keep the row with
the highest value.

```{r remove-duplicate}
# Keep the row with max steps or values
Clean_DailyActivity <- Clean_DailyActivity %>%
  group_by(Id, ActivityDate) %>%
  slice_max(TotalSteps, n = 1, with_ties = FALSE) %>%
  ungroup()
```

```{r check duplicate on DailySleep}
#Checking any duplicate based on user + date on DailySleep dataframe
sum(duplicated(Clean_DailySleep[, c("Id", "SleepDay")]))
```

After running and checking the duplicate on the DailySleep /
Clean_DailySleep dataframe, it appears that there were 3 duplicate rows.
But unlike the Clean_DailyActivity, the duplicate in this dataframe has
the exact value, so I removed the duplicate and only keep the first
occurrence.

```{r remove duplicate on DailySleep}
# Keep only the first occurrence, and remove the rest
Clean_DailySleep <- Clean_DailySleep %>%
  group_by(Id, SleepDay) %>%
  slice(1) %>%
  ungroup()
```

```{r check duplicate on HourlySteps}
# Checking any duplicate based on user + date on Clean_HourlySteps dataframe
sum(duplicated(Clean_HourlySteps[, c("Id", "ActivityHour")]))
```

There are at least 175 duplicates on the HourlySteps / Clean_HourlySteps
dataframe with the same value, just like the previous dataframe. The
next stap is to remove the duplicate and only keep the first occurance.

```{r remove duplicate on HourlySteps}
# Keep only the first occurrence, and remove the rest
Clean_HourlySteps <- Clean_HourlySteps %>%
  group_by(Id, ActivityHour) %>%
  slice(1) %>%
  ungroup()
```

Now, all of the dataframe no longer has any duplicate data.

Next is to figure out if there are missing values in each data frame.

### Check Missing Values

```{r check missing value}
# 1. DailyActivity
colSums(is.na(Clean_DailyActivity))
colMeans(is.na(Clean_DailyActivity)) * 100

# 2. DailySleep
colSums(is.na(Clean_DailySleep))
colMeans(is.na(Clean_DailySleep)) * 100

# 3. HourlySteps
colSums(is.na(Clean_HourlySteps))
colMeans(is.na(Clean_HourlySteps)) * 100
```

Based on this information, there is evidence that no missing values in
each dataframe, and the Data are now clean and ready for analysis.

## **ANALYSIS**

Let's have a look at the summary statistics of the DailyActivity dataset
to find out the overall activity patterns.

## Overall Activity

### Summary 1

```{r Activity Summary}
# Core Activity Summary
Clean_DailyActivity %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes,
         LightlyActiveMinutes,
         FairlyActiveMinutes,
         VeryActiveMinutes,
         Calories) %>%
  summary()
```

### Daily Activity Summary Key Insights

After analyzing 940 daily activity records from 35 users, several key
patterns emerged:

1\. Overall Activity Levels - Users averaged **7,377 steps per day**,
which is below the commonly recommended 10,000 steps. The median of
7,142 steps indicates that half of all recorded days had even fewer
steps, suggesting significant room for improvement in daily activity
levels.

2\. Sedentary Lifestyle Concerns - Perhaps most concerning, users were
sedentary for an average of **16.7 hours per day**. This high sedentary
time is consistent with desk-based work environments but presents health
risks and a key opportunity for intervention.

3\. Exercise Intensity - While users averaged 188 minutes of light
activity daily, they only achieved about **34 minutes of
moderate-to-vigorous physical activity** (fairly + very active minutes
combined). This falls slightly above the WHO minimum recommendation of
21 minutes per day but suggests most users could benefit from more
intense exercise.

Let's make some visualizations based on this

```{r histogram of Daily Steps}
# Histogram of daily steps
ggplot(Clean_DailyActivity, aes(x = TotalSteps)) +
  geom_histogram(binwidth = 1000, fill = "#2E86AB", color = "white") +
  geom_vline(aes(xintercept = mean(TotalSteps)), 
             color = "red", linetype = "dashed", size = 1) +
  labs(title = "Distribution of Daily Steps",
       subtitle = paste("Average:", round(mean(Clean_DailyActivity$TotalSteps)), "steps"),
       x = "Total Steps",
       y = "Frequency") +
  theme_minimal()
```

The distribution reveals significant variability in daily activity. A
notable spike at zero steps indicates days when users either didn't wear
their devices or were completely inactive. The majority of activity days
cluster between 5,000-10,000 steps, with a small group of highly active
days exceeding 20,000 steps. This suggests Bellabeat should focus on
both consistency (reducing zero-step days) and motivation (helping users
reach 10,000 steps more regularly).

```{r histogram of Calories}
# Histogram of calories
ggplot(Clean_DailyActivity, aes(x = Calories)) +
  geom_histogram(binwidth = 200, fill = "#A23B72", color = "white") +
  geom_vline(aes(xintercept = mean(Calories)), 
             color = "red", linetype = "dashed", size = 1) +
  labs(title = "Distribution of Daily Calories Burned",
       subtitle = paste("Average:", round(mean(Clean_DailyActivity$Calories)), "calories"),
       x = "Calories",
       y = "Frequency") +
  theme_minimal()
```

Calories expenditure follows a more normal distribution centered around
2,000-2,500 calories per day, which aligns with typical adult basal
metabolic rates plus light activity. The distribution suggests most
users maintain relatively consistent daily energy expenditure, even when
step counts vary significantly. This indicates that factors beyond steps
(such as body composition and non-tracked activities) play important
roles in total calorie burn.

```{r calculate total active vs sedentary time }
# Calculate total active vs sedentary time
ActivityBreakdown <- Clean_DailyActivity %>%
  summarise(
    Sedentary = mean(SedentaryMinutes),
    `Lightly Active` = mean(LightlyActiveMinutes),
    `Fairly Active` = mean(FairlyActiveMinutes),
    `Very Active` = mean(VeryActiveMinutes)
  ) %>%
  pivot_longer(everything(), names_to = "Activity_Type", values_to = "Minutes")
```

```{r bar chart of Average Daily Activity Minutes by Intensity}
# bar chart
ggplot(ActivityBreakdown, aes(x = reorder(Activity_Type, -Minutes), y = Minutes, fill = Activity_Type)) +
  geom_col() +
  geom_text(aes(label = round(Minutes, 0)), vjust = -0.5) +
  scale_fill_manual(values = c("Sedentary" = "#E63946", 
                                "Lightly Active" = "#F4A261",
                                "Fairly Active" = "#2A9D8F",
                                "Very Active" = "#264653")) +
  labs(title = "Average Daily Activity Minutes by Intensity",
       x = "Activity Level",
       y = "Minutes") +
  theme_minimal() +
  theme(legend.position = "none")
```

This visualization reveals the most concerning finding of our analysis:
Users spend an overwhelming **1,001 minutes (16.7 hours)** per day
sedentary, compared to just 34 minutes of moderate-to-vigorous physical
activity. This 30:1 ratio of sitting to meaningful exercise represents a
significant health risk and a major opportunity for Bellabeat
intervention. While users do achieve 188 minutes of light activity
(likely walking, household tasks), they fall far short of recommended
exercise levels.

```{r box plot}
# Box plot for steps (shows outliers and distribution)
ggplot(Clean_DailyActivity, aes(y = TotalSteps)) +
  geom_boxplot(fill = "#2E86AB", alpha = 0.7) +
  labs(title = "Daily Steps Distribution with Outliers",
       y = "Total Steps") +
  theme_minimal() +
  theme(axis.text.x = element_blank())
```

The box plot confirms the wide variability in user behavior, with 50% of
days falling between 3,300-10,600 steps. Numerous outliers above 20,000
steps suggest users are capable of high activity but don't sustain it
consistently. This indicates Bellabeat should focus on helping users
maintain moderate consistency (7,000-10,000 steps daily) rather than
encouraging occasional extreme activity days.

## User Classification (Sedentary vs. Active)

# Summary 2

### Create user summary

```{r user summary}
user_summary <- Clean_DailyActivity %>%
  group_by(Id) %>%
  summarise(
    avg_steps = mean(TotalSteps),
    avg_calories = mean(Calories),
    avg_active_minutes = mean(VeryActiveMinutes + FairlyActiveMinutes),
    avg_sedentary_minutes = mean(SedentaryMinutes),
    tracking_days = n()
  ) %>%
  ungroup()

# View the data
head(user_summary)
summary(user_summary)
```

```{r categorize}
user_summary <- user_summary %>%
  mutate(
    user_type = case_when(
      avg_steps < 5000 ~ "Sedentary",
      avg_steps < 7500 ~ "Lightly Active",
      avg_steps < 10000 ~ "Fairly Active",
      TRUE ~ "Very Active"
    )
  )

# Show each category with the percentages
user_summary %>%
  count(user_type) %>%
  mutate(percentage = n / sum(n) * 100)
```

Now let's create a visualization based on this

```{r visualization}
user_type_summary <- user_summary %>%
  count(user_type) %>%
  mutate(percentage = n / sum(n) * 100)

ggplot(user_type_summary, aes(x = reorder(user_type, -n), y = n, fill = user_type)) +
  geom_col() +
  geom_text(aes(label = paste0(n, " (", round(percentage, 1), "%)")), 
            vjust = -0.5, size = 4) +
  scale_fill_manual(values = c("Sedentary" = "#E63946", 
                                "Lightly Active" = "#F4A261", 
                                "Fairly Active" = "#2A9D8F", 
                                "Very Active" = "#264653")) +
  labs(title = "User Activity Level Distribution",
       subtitle = "Most users (57%) are sedentary or lightly active",
       x = "Activity Level",
       y = "Number of Users") +
  theme_minimal() +
  theme(legend.position = "none") +
  ylim(0, 13)
```

Interpretation:

1\. **The majority need significant support (57%)**: The largest segment
consist of sedentary (31.4%) and lightly active (25.7%) users, together
representing **57% of the user nase**. These 20 users are not meeting
the CDC's recommended 10.000 daily steps and represent Bellabeat's
target audience for intervention.

2\. **Even distribution across activity levels**: Unlike a typical
population where most people would cluster in one category, our users
show relatively even distribution across all four levels (ranging from
20-31%). This suggests diverse fitness and motivations within the user
base.

3\. **Only 1 in 5 users meets recommended goals**: Just 7 users (20%)
achieve the "very active" classification of 10.000+ steps daily. This
reveals a significant opportunity: **80% of users need help reaching
optimal activity levels**. 4. **The "Almost There" group**: The 8
"Fairly Active" users (22.9%) averaging 7.500 - 10.000 steps represent a
key opportunity—they're close to meeting goals and may respond well to
targeted motivation to push them over te 10.000 step threshold.

## Activity Relationship

### Step vs. Calories Correlation

```{r scatter plot Steps vs. Calories}
# Scatter plot: Steps vs. Calories
ggplot(Clean_DailyActivity, aes(x = TotalSteps, y = Calories)) +
  geom_point(alpha = 0.5, color = "#2E86AB") +
  geom_smooth(method = "lm", color = "#E63946", se = TRUE) +
  labs(title = "Relationship Between Daily Steps and Calories Burned",
       subtitle = paste("Correlation:", 
                       round(cor(Clean_DailyActivity$TotalSteps, Clean_DailyActivity$Calories), 3)),
       x = "Total Steps",
       y = "Calories Burned") +
  theme_minimal()

# Calculate correlation
cor(Clean_DailyActivity$TotalSteps, Clean_DailyActivity$Calories)
```

From this visualization, it is evident that there is a **Strong
Correlation** between the relationship of Daily Steps and Calories
Burned with Correlation Coefficient **0.58**. While steps show a solid
correlation with calories, it's weaker than distance and intensity,
let's take a look further to find out.

### Active Minutes vs. Calories

```{r scatter plot Active Minutes vs. Calories}
# Create total active minutes variable
Clean_DailyActivity <- Clean_DailyActivity %>%
  mutate(TotalActiveMinutes = VeryActiveMinutes + FairlyActiveMinutes + LightlyActiveMinutes)

# Scatter plot: Active Minutes vs. Calories
ggplot(Clean_DailyActivity, aes(x = TotalActiveMinutes, y = Calories)) +
  geom_point(alpha = 0.5, color = "#2A9D8F") +
  geom_smooth(method = "lm", color = "#E63946", se = TRUE) +
  labs(title = "Relationship Between Active Minutes and Calories Burned",
       subtitle = paste("Correlation:", 
                       round(cor(Clean_DailyActivity$TotalActiveMinutes, Clean_DailyActivity$Calories), 3)),
       x = "Total Active Minutes",
       y = "Calories Burned") +
  theme_minimal()

# Calculate correlation
cor(Clean_DailyActivity$TotalActiveMinutes, Clean_DailyActivity$Calories)
```

Next, from this scatter plot visualization, it is evident that there is
a **Moderate Correlation** between Active Minutes and Calories Burned
with Correlation Coefficient **0.469**. Surprisingly, total active
minutes (including light, fair, and very active) shows the weakest
correlation.

### Very Active Minutes vs. Calories

```{r scatter plot Very Active Minutes vs. Calories}
# Scatter plot: Very Active Minutes vs. Calories
ggplot(Clean_DailyActivity, aes(x = VeryActiveMinutes, y = Calories)) +
  geom_point(alpha = 0.5, color = "#F4A261") +
  geom_smooth(method = "lm", color = "#E63946", se = TRUE) +
  labs(title = "Impact of High-Intensity Activity on Calories Burned",
       subtitle = paste("Correlation:", 
                       round(cor(Clean_DailyActivity$VeryActiveMinutes, Clean_DailyActivity$Calories), 3)),
       x = "Very Active Minutes",
       y = "Calories Burned") +
  theme_minimal()

# Calculate correlation
cor(Clean_DailyActivity$VeryActiveMinutes, Clean_DailyActivity$Calories)
```

Moving on to High-Intensity Activity vs. Calories Burned. From this
scatter plot, it is evident that there is a **Strong Correlation**
between High-Intensity and Calories Burned with Correlation Coefficient
**0.594**. Very active minutes show the second-strongest correlation
with calories, despite users averaging only 20 minutes daily of this
intensity.

This reveals an important insight: **quality matters as much as
quantity**.

### Distance vs. Calories

```{r scatter plot Distance vs. Calories}
# Scatter plot: Distance vs. Calories
ggplot(Clean_DailyActivity, aes(x = TotalDistance, y = Calories)) +
  geom_point(alpha = 0.5, color = "#264653") +
  geom_smooth(method = "lm", color = "#E63946", se = TRUE) +
  labs(title = "Relationship Between Distance Traveled and Calories Burned",
       subtitle = paste("Correlation:", 
                       round(cor(Clean_DailyActivity$TotalDistance, Clean_DailyActivity$Calories), 3)),
       x = "Total Distance (km)",
       y = "Calories Burned") +
  theme_minimal()

# Calculate correlation
cor(Clean_DailyActivity$TotalDistance, Clean_DailyActivity$Calories)
```

Last, there is a **Strong Correlation** between Distance and Calories
burned with Coefficient Correlation **0.63**. The relationship between
total distance traveled and calories burned shows the highest
correlation. This makes intuitive sense—covering more ground requires
more energy expenditure regardless of pace.

## Time-Based Pattern

### **Day of Week Analysis** - Are users more active on weekday vs. weekends?

```{r code to add day of week column}
# Add day of week column
Clean_DailyActivity <- Clean_DailyActivity %>%
  mutate(
    day_of_week = wday(ActivityDate, label = TRUE),  # Mon, Tue, Wed...
    day_type = ifelse(day_of_week %in% c("Sat", "Sun"), "Weekend", "Weekday")
  )

# Check it worked
head(Clean_DailyActivity %>% select(ActivityDate, day_of_week, day_type))
```

```{r code to summary stats by day type}
# Summary stats by day type
Clean_DailyActivity %>%
  group_by(day_type) %>%
  summarise(
    avg_steps = mean(TotalSteps),
    avg_calories = mean(Calories),
    avg_distance = mean(TotalDistance),
    avg_active_minutes = mean(VeryActiveMinutes + FairlyActiveMinutes),
    avg_sedentary = mean(SedentaryMinutes)
  )
```

From these results, it is shown that there is **NO Significant
Difference** between Weekday and Weekend. The difference only \~265
steps.

```{r code to calculate average by DOW}
# Calculate average by day of week
daily_summary <- Clean_DailyActivity %>%
  group_by(day_of_week) %>%
  summarise(
    avg_steps = mean(TotalSteps),
    avg_calories = mean(Calories),
    count = n()
  )

# Line chart showing weekly pattern
ggplot(daily_summary, aes(x = day_of_week, y = avg_steps, group = 1)) +
  geom_line(color = "#2E86AB", size = 1.2) +
  geom_point(color = "#2E86AB", size = 3) +
  labs(title = "Average Daily Steps by Day of Week",
       subtitle = "Do users maintain consistency throughout the week?",
       x = "Day of Week",
       y = "Average Steps") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0))
```

```{r code for visualization}
# Box plot comparing weekday vs weekend
ggplot(Clean_DailyActivity, aes(x = day_type, y = TotalSteps, fill = day_type)) +
  geom_boxplot() +
  scale_fill_manual(values = c("Weekday" = "#2E86AB", "Weekend" = "#E63946")) +
  labs(title = "Activity Levels: Weekdays vs Weekends",
       x = "Day Type",
       y = "Total Steps") +
  theme_minimal() +
  theme(legend.position = "none")

# Bar chart with multiple metrics
Clean_DailyActivity %>%
  group_by(day_type) %>%
  summarise(
    Steps = mean(TotalSteps),
    Calories = mean(Calories),
    `Active Minutes` = mean(VeryActiveMinutes + FairlyActiveMinutes)
  ) %>%
  pivot_longer(cols = -day_type, names_to = "Metric", values_to = "Value") %>%
  ggplot(aes(x = day_type, y = Value, fill = day_type)) +
  geom_col() +
  facet_wrap(~Metric, scales = "free_y") +
  scale_fill_manual(values = c("Weekday" = "#2E86AB", "Weekend" = "#E63946")) +
  labs(title = "Weekday vs Weekend Activity Comparison",
       x = "",
       y = "Average Value") +
  theme_minimal() +
  theme(legend.position = "none")
```

### **Hourly Patterns** - When during the day are users most active?

```{r code to convert ActivityHour to datetime and extract hour}
# Step 1: Convert ActivityHour to datetime and extract hour
HourlySteps <- HourlySteps %>%
  mutate(
    ActivityHour = mdy_hms(ActivityHour),  # Convert to datetime
    hour = hour(ActivityHour)               # Extract just the hour (0-23)
  )

# Step 2: NOW create hourly pattern grouped by hour only
hourly_pattern <- HourlySteps %>%
  group_by(hour) %>%  # Group by hour (0-23), not full datetime
  summarise(
    avg_steps = mean(StepTotal),
    median_steps = median(StepTotal),
    total_records = n()
  ) %>%
  arrange(hour)

# View the results
head(hourly_pattern)
```

```{r line chart}
ggplot(hourly_pattern, aes(x = hour, y = avg_steps)) +
  geom_line(color = "#2E86AB", linewidth = 1.2) +
  geom_point(color = "#2E86AB", size = 3) +
  geom_area(alpha = 0.3, fill = "#2E86AB") +
  scale_x_continuous(breaks = seq(0, 23, 2),
                     labels = c("12AM", "2AM", "4AM", "6AM", "8AM", "10AM",
                               "12PM", "2PM", "4PM", "6PM", "8PM", "10PM")) +
  labs(title = "Average Hourly Step Count Throughout the Day",
       subtitle = "Peak activity occurs between 5-7 PM (evening)",
       x = "Hour of Day",
       y = "Average Steps per Hour") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Interpretation:

The hourly activity pattern reveals distinct behavioral trends
throughout the day:

1\. **Sleep Period (12 AM - 5 AM)**: Minimal activity (\~20-50
steps/hour) - Users are clearly asleep during these hours - Baseline
activity from restless movements

2\. **Morning Ramp-Up (6 AM - 8 AM)**: Sharp increase (150 → 400
steps/hour) - Morning routines and commutes begin - Activity doubles
within 2 hours

3\. **Mid-Morning Plateau (9 AM - 11 AM)**: Sustained moderate activity
(\~450-500 steps/hour) - Likely reflects desk work with occasional
movement - Office workers taking short walks

4\. **Lunch Peak (12 PM - 1 PM)**: First daily peak (\~550 steps/hour) -
Lunch break activity - Walking to restaurants or outdoor breaks

5\. **Afternoon Dip (3 PM)**: Notable decrease (\~400 steps/hour) -
Post-lunch slump - "Dead zone" for activity

6\. **Evening Peak (5 PM - 7 PM)**: Highest activity of the day (600+
steps/hour) - **6 PM shows maximum activity** - Post-work exercise,
evening walks - Commute home + intentional fitness

7\. **Evening Wind-Down (8 PM - 11 PM)**: Gradual decline (380 → 200
steps/hour) - Dinner, relaxation, home activities - Preparing for sleep

## Sleep Analysis

### Summary 3

```{r summary code}
# Overall sleep summary
Clean_DailySleep %>%
  mutate(sleep_hours = TotalMinutesAsleep / 60) %>%
  select(TotalMinutesAsleep, sleep_hours, TotalTimeInBed) %>%
  summary()

# More detailed summary
Clean_DailySleep %>%
  mutate(
    sleep_hours = TotalMinutesAsleep / 60,
    time_in_bed_hours = TotalTimeInBed / 60,
    sleep_efficiency = (TotalMinutesAsleep / TotalTimeInBed) * 100
  ) %>%
  summarise(
    avg_sleep_hours = mean(sleep_hours),
    median_sleep_hours = median(sleep_hours),
    min_sleep_hours = min(sleep_hours),
    max_sleep_hours = max(sleep_hours),
    avg_time_in_bed = mean(time_in_bed_hours),
    avg_sleep_efficiency = mean(sleep_efficiency),
    total_sleep_records = n(),
    unique_users = n_distinct(Id)
  )
```

### Sleep Analysis Key Insights

Average Sleep: **6.99 hours** (just below recommended 7 hours). Users
are almost getting enough sleep, but not quite.

```{r code to calculate average sleep per user}
# Calculate average sleep per user
user_sleep_summary <- Clean_DailySleep %>%
  mutate(
    sleep_hours = TotalMinutesAsleep / 60,
    sleep_efficiency = (TotalMinutesAsleep / TotalTimeInBed) * 100
  ) %>%
  group_by(Id) %>%
  summarise(
    avg_sleep_hours = mean(sleep_hours),
    avg_time_in_bed = mean(TotalTimeInBed) / 60,
    avg_sleep_efficiency = mean(sleep_efficiency),
    sleep_records = n()
  ) %>%
  ungroup()

# View the results
head(user_sleep_summary)
summary(user_sleep_summary)
```

Sleep Efficiency: **91.3%** (EXCELLENT!)

\- Above the 85% threshold for good sleep quality

\- Users fall asleep quickly and stay asleep

\- Median of 94% means most users have very efficient sleep

\- Range: 63%-98% shows most users sleep well once in bed

Let's see the visualization based on this information

```{r histogram of Sleep Efficiency}
# Sleep efficiency histogram
Clean_DailySleep %>%
  mutate(sleep_efficiency = (TotalMinutesAsleep / TotalTimeInBed) * 100) %>%
  ggplot(aes(x = sleep_efficiency)) +
  geom_histogram(binwidth = 2, fill = "#1982C4", color = "white") +
  geom_vline(aes(xintercept = 85), color = "red", linetype = "dashed", linewidth = 1) +
  annotate("text", x = 80, y = 40, label = "85% efficiency\nthreshold", color = "red") +
  labs(title = "Sleep Efficiency Distribution",
       subtitle = "Sleep efficiency = (Time Asleep / Time in Bed) × 100%",
       x = "Sleep Efficiency (%)",
       y = "Frequency") +
  theme_minimal()
```

```{r code to classify users by sleep duration}
# Classify users by sleep duration
user_sleep_summary <- user_sleep_summary %>%
  mutate(
    sleep_category = case_when(
      avg_sleep_hours < 6 ~ "Insufficient Sleep (<6h)",
      avg_sleep_hours < 7 ~ "Below Recommended (6-7h)",
      avg_sleep_hours <= 9 ~ "Recommended (7-9h)",
      TRUE ~ "Excessive Sleep (>9h)"
    )
  )

# See distribution
table(user_sleep_summary$sleep_category)

# With percentages
user_sleep_summary %>%
  count(sleep_category) %>%
  mutate(percentage = n / sum(n) * 100) %>%
  arrange(desc(n))
```

Let's create visualization based on this

```{r bar chart of sleep categories}
# Bar chart of sleep categories
user_sleep_summary %>%
  count(sleep_category) %>%
  mutate(
    percentage = n / sum(n) * 100,
    sleep_category = factor(sleep_category, 
                           levels = c("Insufficient Sleep (<6h)",
                                    "Below Recommended (6-7h)",
                                    "Recommended (7-9h)",
                                    "Excessive Sleep (>9h)"))
  ) %>%
  ggplot(aes(x = sleep_category, y = n, fill = sleep_category)) +
  geom_col() +
  geom_text(aes(label = paste0(n, " users\n(", round(percentage, 1), "%)")), 
            vjust = -0.3, size = 4) +
  scale_fill_manual(values = c("Insufficient Sleep (<6h)" = "#E63946",
                                "Below Recommended (6-7h)" = "#F4A261",
                                "Recommended (7-9h)" = "#2A9D8F",
                                "Excessive Sleep (>9h)" = "#264653")) +
  labs(title = "Sleep Quality Classification",
       subtitle = "Based on CDC recommendations (7-9 hours for adults)",
       x = "",
       y = "Number of Users") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 15, hjust = 1)) +
  ylim(0, max(table(user_sleep_summary$sleep_category)) + 3)
```

Interpretation:

\- The largest group with total of 10 users are users with 7-9 hours of
sleep. These individuals meet standard health guidelines for sleep
duration.

\- 'Insufficient Sleep (\<6h)' is the second largest group with total of
8 users. These users are significantly sleep-deprived.

\- 'Below Recommended (6-7h)', these users are slightly under the target
with total of 5 users, often referred to as "short sleepers.

\- A very small minority who sleep longer than the typical clinical
recommendation came from the group with excessive sleep (\>9h), with
only 1 user.

```{r histogram of Sleep Hours}
# Histogram of sleep hours
Clean_DailySleep %>%
  mutate(sleep_hours = TotalMinutesAsleep / 60) %>%
  ggplot(aes(x = sleep_hours)) +
  geom_histogram(binwidth = 0.5, fill = "#6A4C93", color = "white") +
  geom_vline(aes(xintercept = 7), color = "green", linetype = "dashed", linewidth = 1) +
  geom_vline(aes(xintercept = 9), color = "green", linetype = "dashed", linewidth = 1) +
  annotate("text", x = 8, y = 50, label = "Recommended\n7-9 hours", color = "green") +
  labs(title = "Distribution of Sleep Duration",
       subtitle = "Green lines indicate CDC recommended range (7-9 hours)",
       x = "Hours of Sleep",
       y = "Frequency") +
  theme_minimal()
```

## Sleep + Activity Relationship

Next analysis is to find out the relationship between Sleep and overall
activity. But since the data type of *SleepDay* on Clean_DailySleep is
still POSIXct, I need to change the format to date before merging it
with Daily Activity dataframe.

```{r code to convert SleepDay to Date}
# Convert SleepDay from POSIXct to Date
Clean_DailySleep <- Clean_DailySleep %>%
  mutate(SleepDay = as.Date(SleepDay))

# Verify the conversion
class(Clean_DailySleep$SleepDay)
head(Clean_DailySleep$SleepDay)
```

Now the format is set to date, and ready to merge.

```{r merge code}
# Merge Daily Activity and Daily Sleep dataframe
sleep_activity <- Clean_DailyActivity %>%
  inner_join(Clean_DailySleep, by = c("Id" = "Id", "ActivityDate" = "SleepDay"))

# View it
head(sleep_activity)
```

### Summary 4

```{r code to add sleep-related variables}
# Add sleep-related variables
sleep_activity <- sleep_activity %>%
  mutate(
    sleep_hours = TotalMinutesAsleep / 60,
    sleep_efficiency = (TotalMinutesAsleep / TotalTimeInBed) * 100,
    total_active_minutes = VeryActiveMinutes + FairlyActiveMinutes + LightlyActiveMinutes
  )

# Summary
summary(sleep_activity %>% select(sleep_hours, TotalSteps, Calories, total_active_minutes))
```

Next, the relationship I want to find out are:

1\. Is there any significant correlation between Sleep Hour vs. Steps?

2\. Is there any significant correlation between Sleep Hour vs.
Calories? Does sleeps more can burn calories?

3\. Is there any significant correlation between Sleep Hour vs. Active
Minutes?

4\. Is there any significant correlation between Sleep Efficiency vs.
Steps?

```{r code to find out correlation coefficient}
# Key correlations
cor(sleep_activity$sleep_hours, sleep_activity$TotalSteps)
cor(sleep_activity$sleep_hours, sleep_activity$Calories)
cor(sleep_activity$sleep_hours, sleep_activity$total_active_minutes)
cor(sleep_activity$sleep_efficiency, sleep_activity$TotalSteps)

# Create a correlation summary
sleep_correlations <- data.frame(
  Metric = c("Sleep Hours vs Steps", 
             "Sleep Hours vs Calories",
             "Sleep Hours vs Active Minutes",
             "Sleep Efficiency vs Steps"),
  Correlation = c(
    cor(sleep_activity$sleep_hours, sleep_activity$TotalSteps),
    cor(sleep_activity$sleep_hours, sleep_activity$Calories),
    cor(sleep_activity$sleep_hours, sleep_activity$total_active_minutes),
    cor(sleep_activity$sleep_efficiency, sleep_activity$TotalSteps)
  )
)

print(sleep_correlations)
```

Before finally revealing the interpretation behind these results, let's
create a visualization based on this

```{r scatter plot}
# Visualization 1: Sleep vs Steps
ggplot(sleep_activity, aes(x = sleep_hours, y = TotalSteps)) +
  geom_point(alpha = 0.5, color = "#6A4C93") +
  geom_smooth(method = "lm", color = "#E63946", se = TRUE) +
  geom_vline(xintercept = 7, linetype = "dashed", color = "green", alpha = 0.5) +
  geom_vline(xintercept = 9, linetype = "dashed", color = "green", alpha = 0.5) +
  annotate("text", x = 8, y = max(sleep_activity$TotalSteps) * 0.9, 
           label = "Recommended\n7-9 hours", color = "green", size = 3) +
  labs(title = "Sleep Duration vs Daily Steps: No Positive Relationship",
       subtitle = "Correlation: -0.19 (weak negative)",
       x = "Hours of Sleep",
       y = "Total Steps") +
  theme_minimal()
```

The Surprising Finding: **No Positive Correlation**

Contrary to conventional wisdom that "better sleep leads to more
activity," this analysis reveals **no positive relationship** between
sleep duration and physical activity levels.

**Correlation Analysis:**

\- Sleep Hours vs Steps: **-0.19** (weak negative)

\- Sleep Hours vs Calories: **-0.03** (essentially zero)

\- Sleep Hours vs Active Minutes: **-0.07** (essentially zero)

\- Sleep Efficiency vs Steps: **-0.11** (weak negative)

The scatter plot visualization clearly shows a **slight downward
trend** - this means users who sleep more tend to take slightly fewer
steps, though the relationship is weak and highly variable.

Next, let's find out the relationship between Daily Steps and Sleep
Quality.

```{r code to calculate average steps by sleep category}
# Average steps by sleep category
sleep_activity <- sleep_activity %>%
  mutate(
    sleep_category = case_when(
      sleep_hours < 6 ~ "Insufficient (<6h)",
      sleep_hours < 7 ~ "Below Rec. (6-7h)",
      sleep_hours <= 9 ~ "Recommended (7-9h)",
      TRUE ~ "Excessive (>9h)"
    ),
    sleep_category = factor(sleep_category,
                           levels = c("Insufficient (<6h)",
                                    "Below Rec. (6-7h)",
                                    "Recommended (7-9h)",
                                    "Excessive (>9h)"))
  )

# Bar chart
sleep_activity %>%
  group_by(sleep_category) %>%
  summarise(
    avg_steps = mean(TotalSteps),
    count = n()
  ) %>%
  ggplot(aes(x = sleep_category, y = avg_steps, fill = sleep_category)) +
  geom_col() +
  geom_text(aes(label = paste0(round(avg_steps, 0), " steps\n(n=", count, ")")),
            vjust = -0.3, size = 3.5) +
  scale_fill_manual(values = c("Insufficient (<6h)" = "#E63946",
                                "Below Rec. (6-7h)" = "#F4A261",
                                "Recommended (7-9h)" = "#2A9D8F",
                                "Excessive (>9h)" = "#264653")) +
  labs(title = "Average Daily Steps by Sleep Quality",
       subtitle = "Well-rested users don't necessarily take more steps",
       x = "",
       y = "Average Steps") +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 15, hjust = 1)) +
  ylim(0, max(tapply(sleep_activity$TotalSteps, sleep_activity$sleep_category, mean)) * 1.15)
```

## Device Usage & Tracking Consistency

This analysis is to understanding how consistently users engage with
their fitness trackers reveals critical insights about user behavior and
product stickiness.

### Activity Tracking

```{r code to count how many days did each user track activity}
# How many days did each user track activity?
user_tracking <- Clean_DailyActivity %>%
  group_by(Id) %>%
  summarise(
    tracking_days = n(),
    avg_steps = mean(TotalSteps)
  ) %>%
  arrange(desc(tracking_days))

# Summary statistics
summary(user_tracking$tracking_days)

# View the data
head(user_tracking)
```

Let's create a visualization based on this summary

```{r viz code for Activity Tracking Distribution}
# Viz1: Activity Tracking Distribution
ggplot(user_tracking, aes(x = tracking_days)) +
  geom_histogram(binwidth = 5, fill = "#2E86AB", color = "white") +
  geom_vline(aes(xintercept = median(tracking_days)), 
             color = "red", linetype = "dashed", linewidth = 1) +
  annotate("text", x = median(user_tracking$tracking_days) + 5, y = 8,
           label = paste("Median:", median(user_tracking$tracking_days), "days"),
           color = "red") +
  labs(title = "User Engagement: Days of Activity Tracking",
       subtitle = "How consistently do users wear their devices?",
       x = "Number of Days Tracked",
       y = "Number of Users") +
  theme_minimal()
```

Activity Tracking: Strong Engagement

**Finding**: Users demonstrate solid commitment to activity tracking,
with a median of **42 days** tracked over the study period.

**Key Statistics**:

\- Median: 42 days

\- Average: 39.2 days

\- Range: 8-62 days

\- **Most users (20 out of 35)** tracked for 38-42+ days

**Interpretation**: The concentration of users around the 40-day mark
suggests good device adoption and habit formation. Users who start
tracking tend to stick with it for at least a month, indicating the
value proposition for activity tracking is clear and compelling.

### Sleep Tracking

```{r code to count how many days did each user track sleep}
# How many days did each user track sleep?
sleep_tracking <- Clean_DailySleep %>%
  group_by(Id) %>%
  summarise(sleep_days = n()) %>%
  arrange(desc(sleep_days))

# Summary statistics
summary(sleep_tracking$sleep_days)

# View it
head(sleep_tracking)
```

```{r code to merge tracking data}
# Merge tracking data
tracking_comparison <- user_tracking %>%
  left_join(sleep_tracking, by = "Id") %>%
  mutate(sleep_days = ifelse(is.na(sleep_days), 0, sleep_days))

# Summary
tracking_comparison %>%
  summarise(
    avg_activity_days = mean(tracking_days),
    avg_sleep_days = mean(sleep_days, na.rm = TRUE),
    users_tracking_sleep = sum(sleep_days > 0)
  )
```

Sleep Tracking: The Engagement Gap

**Finding**: Sleep tracking shows significantly lower engagement, with a
median of only **20.5 days** - less than half the activity tracking
rate.

**Key Statistics**:

\- Median: 20.5 days (among those who tracked)

\- Average across all users: 11.7 days

\- **31% of users (11 out of 35) never tracked sleep at all**

\- **Sleep tracking is 50% less consistent** than activity tracking

```{r viz code for Activity vs. Sleep Tracking Comparison}
# Viz2: Activity vs. Sleep Tracking Comparison
tracking_long <- tracking_comparison %>%
  select(Id, tracking_days, sleep_days) %>%
  pivot_longer(cols = c(tracking_days, sleep_days),
               names_to = "tracking_type",
               values_to = "days") %>%
  mutate(tracking_type = ifelse(tracking_type == "tracking_days", 
                                "Activity Tracking", 
                                "Sleep Tracking"))

# Box plot comparison
ggplot(tracking_long, aes(x = tracking_type, y = days, fill = tracking_type)) +
  geom_boxplot() +
  scale_fill_manual(values = c("Activity Tracking" = "#2E86AB", 
                                "Sleep Tracking" = "#6A4C93")) +
  labs(title = "Tracking Consistency: Activity vs Sleep",
       subtitle = "Sleep tracking is significantly less consistent",
       x = "",
       y = "Days Tracked") +
  theme_minimal() +
  theme(legend.position = "none")
```

**The Dramatic Difference**: The box plot visualization clearly shows
activity tracking clustered around 40 days, while sleep tracking has a
much wider distribution with many users at zero or very low numbers.

### Consistency vs. Activity Level

```{r viz ciode}
# Does tracking consistency correlate with activity?
ggplot(user_tracking, aes(x = tracking_days, y = avg_steps)) +
  geom_point(color = "#2E86AB", size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", color = "#E63946", se = TRUE) +
  labs(title = "Does Consistent Tracking Lead to More Steps?",
       subtitle = paste("Correlation:", 
                       round(cor(user_tracking$tracking_days, user_tracking$avg_steps), 3)),
       x = "Days Tracked",
       y = "Average Daily Steps") +
  theme_minimal()

# Calculate correlation
cor(user_tracking$tracking_days, user_tracking$avg_steps)
```

Engagement Drives Results

**Critical Finding**: Tracking consistency correlates positively with
activity levels (r = 0.33).

Users who tracked for **40+ days averaged \~8,000 steps**, while those
tracking **fewer than 20 days averaged \~4,000 steps**.

This suggests:

\- Consistent tracking creates accountability

\- Regular data viewing motivates behavior change

\- Habit formation requires sustained engagement

## **EXECUTIVE SUMMARY & ACTIONABLE RECOMMENDATIONS**

Bellabeat Case Study: Key findings & strategic recommendations

## Executive Summary

Analysis of 35 Fitbit users' activity, sleep, and usage patterns reveals
critical opportunities for Bellabeat to differentiate itself in the
wellness technology market. My findings challenge conventional wisdom
about sleep-activity relationships while identifying clear user segments
and engagement gaps.

------------------------------------------------------------------------

## Top 5 Critical Findings

1.  The Sedentary Crisis (16.7 hours/day) Users are sedentary for over
    16 hours daily but achieve only 34 minutes of moderate-to-vigorous
    activity. This represents the **primary health risk** and **biggest
    intervention opportunity**.

2.  Majority Users Need Support (57%) 31% are sedentary and 26% are
    lightly active - together representing the core target audience who
    need motivation and behavioral change support.

3.  Sleep-Activity Paradox Better sleep does NOT lead to more activity
    (r = -0.19). Sleep and activity are **independent behaviors**
    requiring separate interventions, not a simple "sleep more → move
    more" message.

4.  Sleep Tracking Engagement Gap Only 68% of users track sleep, and
    those who do track **50% less consistently** than activity (20 vs 42
    days median). This represents a major UX challenge.

5.  Engagement Drives Results Consistent tracking correlates with higher
    activity (r = 0.33). Users tracking 40+ days average 8,000 steps vs
    4,000 steps for those tracking \<20 days.

------------------------------------------------------------------------

## Strategic Recommendations for Bellabeat

**PRIORITY 1: Combat Sedentary Behavior**

**The Problem**: 16.7 hours/day sedentary time is unsustainable and
unhealthy.

**Solutions**:

-   Giving **Hourly Movement Reminders** for users within this category:
    Example of reminder, "You've been sitting for 50 minutes - time for
    a 2-minute walk!"
-   Showing **Sedentary Time Tracking**: Make sitting visible like
    steps, for example, "You sat for 14 hours today. Get up and move!"
-   Set **Micro-Movement Goals**: "Take 250 steps every hour" instead of
    overwhelming 10,000-step goals
-   **3 PM Wake-Up Call**: Target the afternoon activity dip with
    energizing challenges

**PRIORITY 2: Segment-Specific Engagement**

**The Problem**: 57% of users are sedentary or lightly active and need
different support than the 20% who are very active.

So here's the recommendation for every category:

1.  **Sedentary Users (31%)** - "The Beginners"

-   Start with 3,000-5,000 step goals
-   Celebrate ANY increase, for example, "You walked 500 more steps than
    yesterday!"
-   Focus on consistency over volume
-   Gentle, encouraging tone

2.  **Lightly Active Users (26%)** - "The Strivers"

-   Progressive goals: 5K → 7.5K → 10K steps
-   Weekly challenges with achievable targets
-   Social features for accountability
-   Highlight health benefits of incremental improvements

3.  **Fairly Active Users (23%)** - "The Almost-Theres"

-   Push toward 10K steps with targeted motivation
-   Introduce intensity tracking (not just steps)
-   Competitive leaderboards
-   Reward consistency streaks

4.  **Very Active Users (20%)** - "The Athletes"

-   Advanced analytics (VO2 max, recovery metrics)
-   Training plans and periodization
-   Performance optimization insights
-   Community mentorship opportunities

**PRIORITY 3: Fix Sleep Tracking UX**

**The Problem**: 31% never track sleep; those who do track 50% less than
activity.

**Solutions**:

1.  **Auto-Sleep Detection**: Remove the need to manually activate sleep
    mode

2.  **Charging Solutions**:

    -   Rapid charging (80% in 30 minutes during morning routine)

    -   Alternative: Two devices (wear one while charging the other)

3.  **Comfort First**: Smaller, lighter sleep-specific device or
    redesigned band

4.  **Demonstrate Value**: Show "Sleep Score" and next-day energy
    predictions immediately

5.  **Bedtime Reminders**: "To get 8 hours, go to bed by 10:30 PM"

**PRIORITY 4: Optimal Timing for Interventions**

**The Problem**: Users have distinct activity patterns that current
generic notifications ignore.

**Time-Based Strategy**:

-   **5-7 PM (Peak Activity)**: Send workout challenges, achievement
    celebrations
-   **12-1 PM (Lunch Peak)**: "Take a 10-minute walk after lunch"
-   **3 PM (Afternoon Dip)**: "Energy slump? 5-minute movement break!"
-   **11:30 AM**: Pre-lunch walk reminders
-   **9 PM**: Wind-down routine prompts for better sleep
-   **Avoid 8-10 AM**: Commute chaos, low engagement window

**PRIORITY 5: Emphasize Intensity Over Volume**

**The Problem**: Users fixate on steps but ignore intensity, despite
intensity having stronger correlation with calories (r = 0.59 vs 0.58).

**Solutions**:

\- **Intensity Zones**: Display time in heart rate zones prominently

\- **Quality Over Quantity**: "20 minutes of vigorous activity \> 20,000
light steps"

\- **10-Minute HIIT Challenges**: For time-constrained users

\- **Distance + Intensity Metrics**: Track both steps and "active
distance"

\- **Reframe Success**: "You hit your intensity goal!" not just "You hit
your steps!"

**PRIORITY 6: Build Habit Formation Features**

**The Problem**: Tracking consistency predicts success, but many users
don't build lasting habits.

**Solutions**:

-   **Onboarding Excellence**: First 7 days are critical - daily
    check-ins and wins
-   **Streak Tracking**: "You've tracked 15 days in a row! Keep it
    going!"
-   **30-Day Milestone**: Major celebration and reward at 1 month
-   **Habit Stacking**: "Track your morning steps right after your
    coffee"
-   **Social Accountability**: Share streaks with friends
-   **Recovery Paths**: "You missed yesterday - let's get back on track
    today"

**PRIORITY 7: Separate Sleep & Activity Messaging**

**The Problem**: No positive correlation between sleep and activity
challenges the "sleep better → move more" narrative.

**New Approach**:

-   **Independent Value Props**: Promote each for its own benefits
-   **No False Promises**: Don't claim sleep will directly increase
    activity
-   **Holistic Wellness**: "Great sleep AND great activity = optimal
    health"
-   **Different Motivations**: Some users prioritize sleep, others
    prioritize fitness - honor both
-   **Targeted Messaging**: - Active + Poor Sleep: "Protect your
    performance with recovery" - Sedentary + Good Sleep: "You're
    well-rested - let's move!"

------------------------------------------------------------------------

## Conclusion

The path to better health isn't about revolutionary changes - it's about
**sustainable, personalized, and consistent small improvements**.
Bellabeat's opportunity lies in meeting users where they are,
understanding their unique challenges, and providing intelligent,
compassionate support that respects the complexity of real human
behavior.

**The winning formula**: Segment intelligently + Reduce friction + Time
interventions + Celebrate progress + Build habits.

That's it from me for the Capstone project in the Bellabeat Case Study.
Thank you so much for your interest in the project!

**Alifia Ganjaraharja**
